Legion: Expressing Locality and Independence with Logical Regions

### 2012-05-26 doc/asplos2013/legion-llr.tex

\documentclass[9pt,nocopyrightspace,preprint]{sigplanconf}

\bibliographystyle{plainnat}
\usepackage{amsmath,amsthm}
\usepackage{listings}
\usepackage{tikz}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{cancel}

@@

\title{Legion: Expressing Locality and Independence with Logical Regions}
\authorinfo{}{}{}
\maketitle

\begin{abstract}
Limit is 12 pages, including bibliography and appendices.
Deadlines: \\

\begin{tabular}{ll}
Paper registration & July 7, 2012 3:59am PDT \\
Paper submission & July 11, 2012 3:59am PDT \\
Author response & September 10 - September 13, 2012 \\
Author notification & October 1, 2012
\end{tabular}
\end{abstract}

>>>>

\documentclass{sig-alternate}

\bibliographystyle{plainnat}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{listings}
\usepackage{tikz}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{cancel}

@@

\title{Legion: Expressing Locality and Independence with Logical Regions}
\numberofauthors{3}
\author{}
\maketitle

\begin{abstract}
Last year's limit was 12 pages, including bibliography and appendices.
Deadlines:
\begin{tabular}{ll}
Abstract deadline & July 16, 2012 3pm PDT \\
Paper deadline & July 23, 2012 3pm PDT \\
Author response & around October 15, 2012 \\
Author notification & November 9, 2012
\end{tabular}
\end{abstract}

### 2012-05-26 doc/sc2012/circuit.tex

The  {\tt calc\_new\_currents} task reads and writes the wires subregion 
and reads the private, shared, and ghost node subregions for its piece.
The {\tt distribute\_charge} task reads the piece's 
wires subregion and updates all nodes those wires touch.  However,
rather than using read/write permission for the nodes (which would
serialize these tasks for correctness), the task
uses reorderable reduction operations and 
atomic rather than exclusive access. The final task 
{\tt update\_voltages} writes the shared and private nodes for a piece
and reads the results of the previous task's reductions.

>>>>

The  {\tt calc\_new\_currents} task reads and writes the wires subregion 
and reads the private, shared, and ghost node subregions for its piece.
The {\tt distribute\_charge} task reads the piece's 
wires subregion and updates all nodes those wires touch.  However,
rather than using read/write privilege for the nodes (which would
serialize these tasks for correctness), the task
uses reorderable reduction operations and 
atomic rather than exclusive access. The final task 
{\tt update\_voltages} writes the shared and private nodes for a piece
and reads the results of the previous task's reductions.

### 2012-05-26 doc/sc2012/code_ex.tex

void simulate_circuit(Circuit c, float dt) : RWE(c.r_all_nodes,c.r_all_wires)
{
  // The construction of the colorings is not shown.  The colorings wire_owner_map,
  // node_owner_map, and node_neighbor_map have MAX_PIECES colors 
  // 0..MAX_PIECES #$-$# 1. The coloring node_sharing map has two colors 0 and 1.
  //
  // Partition of wires into MAX_PIECES pieces
  partition<disjoint> p_wires = c.r_all_wires.partition(wire_owner_map); 
  // Partition nodes into two parts for all-private vs. all-shared
  partition<disjoint> p_nodes_pvs = c.r_all_nodes.partition(node_sharing map);

@@

void calc_new_currents(CircuitPiece piece):
        RWE(piece.rw_pvt), ROE(piece.rn_pvt,piece.rn_shr,piece.rn_ghost) {
  foreach(w : piece.rw_pvt)
    w#$\rightarrow$#current = (w#$\rightarrow$#in_node#$\rightarrow$#voltage - w#$\rightarrow$#out_node#$\rightarrow$#voltage) / w#$\rightarrow$#resistance;
}

void distribute_charge(CircuitPiece piece, float dt):
        ROE(piece.rw_pvt), RdA(piece.rn_pvt,piece.rn_shr,piece.rn_ghost) {
  foreach(w : piece.rw_pvt) {
    w#$\rightarrow$#in_node#$\rightarrow$#new_charge += -dt * w#$\rightarrow$#current;
    w#$\rightarrow$#out_node#$\rightarrow$#new_charge +=  dt * w#$\rightarrow$#current;
  }
}

void update_voltages(CircuitPiece piece): RWE(piece.rn_pvt,piece.rn_shr) {
  foreach(n : piece.rn_pvt, piece.rn_shr) {
    n#$\rightarrow$#voltage += n#$\rightarrow$#new_charge / n#$\rightarrow$#capacitance;
    n#$\rightarrow$#new_charge = 0;
  }
}
\end{lstlisting}

>>>>

void simulate_circuit(Circuit c, float dt) : RWE(c.r_all_nodes, c.r_all_wires)
{
  // The construction of the colorings is not shown.  The colorings wire_owner_map,
  // node_owner_map, and node_neighbor_map have MAX_PIECES colors 
  // 0..MAX_PIECES #$-$# 1. The coloring node_sharing map has two colors 0 and 1.
  //
  // Partition of wires into MAX_PIECES pieces
  partition<disjoint> p_wires = c.r_all_wires.partition(wire_owner_map); 
  // Partition nodes into two parts for all-private vs. all-shared
  partition<disjoint> p_nodes_pvs = c.r_all_nodes.partition(node_sharing map);

@@

void calc_new_currents(CircuitPiece piece):
        RWE(piece.rw_pvt), ROE(piece.rn_pvt, piece.rn_shr, piece.rn_ghost) {
  foreach(w : piece.rw_pvt)
    w#$\rightarrow$#current = (w#$\rightarrow$#in_node#$\rightarrow$#voltage - w#$\rightarrow$#out_node#$\rightarrow$#voltage) / w#$\rightarrow$#resistance;
}

void distribute_charge(CircuitPiece piece, float dt):
        ROE(piece.rw_pvt), RdA(piece.rn_pvt, piece.rn_shr, piece.rn_ghost) {
  foreach(w : piece.rw_pvt) {
    w#$\rightarrow$#in_node#$\rightarrow$#new_charge += -dt * w#$\rightarrow$#current;
    w#$\rightarrow$#out_node#$\rightarrow$#new_charge +=  dt * w#$\rightarrow$#current;
  }
}

void update_voltages(CircuitPiece piece): RWE(piece.rn_pvt, piece.rn_shr) {
  foreach(n : piece.rn_pvt, piece.rn_shr) {
    n#$\rightarrow$#voltage += n#$\rightarrow$#new_charge / n#$\rightarrow$#capacitance;
    n#$\rightarrow$#new_charge = 0;
  }
}
\end{lstlisting}

### 2012-05-26 doc/sc2012/experiments.tex

Our second experiment is a port of the \emph{fluidanimate} benchmark
from the PARSEC benchmark suite\cite{bienia11benchmarking}, which does a
particle-based simulation of an incompressible fluid.  Each particle
interacts only with nearby particles. The benchmark divides the
space in which the fluid can move into a three-dimensional array of
cells such that the range of interaction is limited to just the cells
adjacent (including diagonals) to the one a particle resides in.  The
application divides the array into {\em grids} and assigns each grid to
a thread.  Per-particle locking safely accesses particles in
cells that lie on the edge of a grid.  This fine-grained locking
scheme along with the assumption of a shared address space gives
good scaling in a multi-core processor, but prohibits running
beyond a single node.

>>>>

Our second experiment is a port of the \emph{fluidanimate} benchmark
from the PARSEC benchmark suite\cite{bienia11benchmarking}, which does a
particle-based simulation of an incompressible fluid.  Each particle
interacts only with nearby particles. The benchmark divides the
space in which the fluid can move into a three-dimensional array of
cells such that the range of interaction is limited to just the cells
adjacent (including diagonals) to the one a particle resides in.  The
application divides the array into {\em grids} and assigns each grid to
a thread.  Per-particle locking safely accesses particles in
cells that lie on the edge of a grid.  This fine-grained locking
scheme along with the assumption of a shared address space gives
good scaling in a multicore processor, but prohibits running
beyond a single node.

### 2012-05-26 doc/sc2012/mapping.tex

The mapping interface consists of ten methods that SOOPs call for
mapping decisions.  A {\em mapper} implementing these methods has access to a simple interface
for inspecting properties of the machine, including a list of
processors and their type (i.e. CPU, GPU) and a list of
memories visible to each processor and their latencies and bandwidths.
For brevity we only discuss three of the mapping interface calls:

@@

The mapping interface has two desirable properties.  First, all mapper
decisions are orthogonal to program correctness and can only impact
performance.  Regardless of where a mapper places a task or region,
the SOOPs schedule tasks and copies in accordance with the privileges
and coherence properties specified in the program.  Therefore when
writing a Legion application, a programmer can begin by using the
default mapper and later improve performance by creating and refining a custom mapper.
Second, the mapping interface isolates machine-specific decisions
to the mapper.  As a result, Legion programs are highly
portable.  To port a Legion program to a new architecture, a programmer need only
implement a new mapper with decisions specific to the new architecture.

@@

Each custom mapper extends the default mapper.  A programmer need only override the functions necessary 
for creating a custom mapping.  Custom mappers can be used to create totally static mappings, 
mappings that memoize their results, or even totally dynamic mappings.  We describe examples of 
custom mappers in Section \ref{sec:exp}.

>>>>

The mapping interface consists of ten methods that SOOPs call for
mapping decisions.  A {\em mapper} implementing these methods has access to a simple interface
for inspecting properties of the machine, including a list of
processors and their type (i.e. CPU, GPU), a list of
memories visible to each processor, and their latencies and bandwidths.
For brevity we only discuss the three most important interface calls:

@@

The mapping interface has two desirable properties.  
First, program correctness is unaffected by mapper decisions, which can only impact 
%First, all mapper
%decisions are orthogonal to program correctness and can only impact
performance.  Regardless of where a mapper places a task or region,
the SOOPs schedule tasks and copies in accordance with the privileges
and coherence properties specified in the program.  Therefore, when
writing a Legion application, a programmer can begin by using the
default mapper and later improve performance by creating and refining a custom mapper.
Second, the mapping interface isolates machine-specific decisions
to the mapper.  As a result, Legion programs are highly
portable.  To port a Legion program to a new architecture, a programmer need only
implement a new mapper with decisions specific to the new architecture.

@@

Each custom mapper extends the default mapper.  A programmer need only override the functions 
%necessary for creating a custom mapping.  
he wishes to customize.  Custom mappers can be used to create totally static mappings, 
mappings that memoize their results, or even totally dynamic mappings.  We describe examples of 
custom mappers in Section \ref{sec:exp}.
